### 2025-03-08 README.qmd
Fully pivoting to the new repo structure

---
title: "The mzsql project"
format: gfm
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

This repository implements "vanilla" SQL databases for the efficient storage and access of raw mass-spectrometry (MS) data. Here, "raw" MS data refers to the data encoded in vendor or .mzML files and contains coordinates of retention time, m/z ratio, and intensity.

Existing data formats fail to provide intuitive, rapid, and programmatic search of raw MS data and require learning the quirks and conceits of idiosyncratic file formats. Databases have consistent and transferrable syntax for queries, indexing and binary search for rapid data extraction, benefit from multi-file aggregations of data, can store processed data alongside, and are established as an industry standard.

## Convert mzML files into a database via the converters above

Converters from mzML to SQLite and DuckDB are currently available in the [converters folder](converters) for R and Python with Julia and C++ hopefully soon. Code from the associated script can be directly copied and pasted into your workflow, though you're responsible for managing the dependency installations.

The database format is shown in the graphic below:

### Convert from mzML to DuckDB using R:

Ensure `RaMS`, `DBI`, and `duckdb`/`RSQLite` are all installed from CRAN.

```{r make database, eval=FALSE}
library(RaMS)
msdata <- grabMSdata(list.files("demo_data", pattern = "mzML", full.names = TRUE))

library(DBI)
conn <- dbConnect(RSQLite::SQLite(), "demo_data/msdata.sqlite")
# conn <- dbConnect(duckdb::duckdb(), "demo_data/msdata.duckdb")
dbWriteTable(conn, "MS1", msdata$MS1)
dbWriteTable(conn, "MS2", msdata$MS2)
dbDisconnect(conn)
```

## Access the database using SQL

SQL is increasingly popular among mass spectrometrists and is consistent between programming languages.

### Extract a chromatogram

General query: `SELECT * FROM MS1 WHERE mz BETWEEN mzmin AND mzmax`

<details open>
<summary>R code</summary>

Ensure `DBI`, and `RSQLite`/`duckdb` are all installed from CRAN.

```{r rchrom}
library(DBI)
pmppm <- function(mass, ppm)c(mass * (1 - ppm/1e+06), mass * (1 + ppm/1e+06))

conn <- dbConnect(RSQLite::SQLite(), "demo_data/msdata.sqlite")
# conn <- dbConnect(duckdb::duckdb(), "demo_data/msdata.duckdb")
chr_statement <- "SELECT * FROM MS1 WHERE mz BETWEEN ? AND ?"
chrom <- dbGetQuery(conn, chr_statement, params=pmppm(118.0865, 10))
dbDisconnect(conn)
plot(chrom$rt, chrom$int, type = "l")
```

</details>

<details>
<summary>Python</summary>

Make sure the duckdb/sqlite and matplotlib packages are installed.

```{python pychrom}
import duckdb
def pmppm(mz, ppm):
  return(mz*(1-ppm/1000000), mz*(1+ppm/1000000))

conn = duckdb.connect("demo_data/msdata.duckdb")
chr_statement = "SELECT * FROM MS1 WHERE mz BETWEEN ? AND ?"
chrom = conn.execute(chr_statement, pmppm(118.0865, 10)).fetchdf()
conn.close()

import matplotlib.pyplot as plt
plt.plot(chrom["rt"], chrom["int"])
plt.show()
```

</details>

<details>
<summary>Julia</summary>

Make sure the SQLite/DuckDB, DataFrames, DBInterface, and Plots packages are installed.

```{julia jchrom}
using DuckDB, DataFrames, Plots;
pmppm(mass, ppm) = [mass * (1 - ppm / 1e6), mass * (1 + ppm / 1e6)];

conn = DBInterface.connect(DuckDB.DB, "demo_data/msdata.duckdb");
chr_statement = "SELECT * FROM MS1 WHERE mz BETWEEN ? AND ?";
chrom = DataFrame(DBInterface.execute(conn, chr_statement, pmppm(118.0865, 10)));
DBInterface.close!(conn)

plot(chrom.rt, chrom.int, lw=2)
```

</details>

### Search for fragments of a specific precursor

General query: `SELECT * FROM MS2 WHERE premz BETWEEN mzmin AND mzmax`

<details open>
<summary>R code</summary>

Ensure `DBI`, `ggplot2`, and `RSQLite`/`duckdb` are all installed from CRAN.

```{r rpre}
library(DBI)
pmppm <- function(mass, ppm)c(mass * (1 - ppm/1e+06), mass * (1 + ppm/1e+06))

conn <- dbConnect(RSQLite::SQLite(), "demo_data/msdata.sqlite")
# conn <- dbConnect(duckdb::duckdb(), "demo_data/msdata.duckdb")
frag_statement <- "SELECT * FROM MS2 WHERE premz BETWEEN ? AND ?"
frag_data <- dbGetQuery(conn, frag_statement, params=pmppm(118.0865, 10))
dbDisconnect(conn)

library(ggplot2)
ggplot(frag_data, aes(x=fragmz, y=int)) +
  geom_point() +
  geom_segment(aes(xend=fragmz, yend=0)) +
  facet_wrap(~filename, ncol=1) +
  scale_y_log10()
```

</details>

<details>
<summary>Python</summary>

Make sure the duckdb and matplotlib packages are installed.

```{python pypre}
import duckdb
import matplotlib.pyplot as plt
def pmppm(mz, ppm):
  return(mz*(1-ppm/1000000), mz*(1+ppm/1000000))

conn = duckdb.connect("demo_data/msdata.duckdb")
frag_statement = "SELECT * FROM MS2 WHERE premz BETWEEN ? AND ?"
frag_data = conn.execute(frag_statement, pmppm(118.0865, 10)).fetchdf()
fig, axes = plt.subplots(3, 1, sharex=True)
for ax, filename in zip(axes, frag_data["filename"].unique()):
    subset = frag_data[frag_data["filename"] == filename]
    ax.vlines(subset["fragmz"], 0, subset["int"], color='k', zorder=1, linewidths=0.5)
    ax.scatter(subset["fragmz"], subset["int"], color='k', zorder=2, s=10)
    ax.set_yscale("log")
    ax.set_title(filename)
conn.close()

plt.tight_layout()
plt.show()
```

</details>

<details>
<summary>Julia</summary>

Make sure the SQLite/DuckDB, DataFrames, DBInterface, and StatsPlots packages are installed.

```{julia jpre}
using DuckDB, DataFrames, StatsPlots;
pmppm(mass, ppm) = [mass * (1 - ppm / 1e6), mass * (1 + ppm / 1e6)];

conn = DBInterface.connect(DuckDB.DB, "demo_data/msdata.duckdb");
frag_statement = "SELECT * FROM MS2 WHERE premz BETWEEN ? AND ?";
frag_data = DataFrame(DBInterface.execute(conn, frag_statement, pmppm(118.0865, 10)));
DBInterface.close!(conn)

@df frag_data scatter(:fragmz, :int, markersize=3, group=:filename, yscale=:log10)
# @df frag_data plot!([:fragmz :fragmz]', [zeros(size(frag_data, 1)) frag_data.int]', lw=1, group=:filename)
```

</details>

### Search for precursors of a specific fragment

General query: `SELECT * FROM MS2 WHERE fragmz BETWEEN mzmin AND mzmax`

<details open>
<summary>R code</summary>

Ensure `DBI`, `ggplot2`, and `duckdb` are all installed from CRAN.

```{r rfrag}
library(DBI)
pmppm <- function(mass, ppm)c(mass * (1 - ppm/1e+06), mass * (1 + ppm/1e+06))

conn <- dbConnect(RSQLite::SQLite(), "demo_data/msdata.sqlite")
# conn <- dbConnect(duckdb::duckdb(), "demo_data/msdata.duckdb")
pre_statement <- "SELECT * FROM MS2 WHERE fragmz BETWEEN ? AND ?"
pre_data <- dbGetQuery(conn, pre_statement, params=pmppm(59.07336, 10))
dbDisconnect(conn)

library(ggplot2)
ggplot(pre_data, aes(x=premz, y=int)) +
  geom_point() +
  geom_segment(aes(xend=premz, yend=0)) +
  facet_wrap(~filename, ncol=1)
```

</details>

<details>
<summary>Python</summary>

Make sure the duckdb and matplotlib packages are installed.

```{python pyfrag}
import duckdb
import matplotlib.pyplot as plt
def pmppm(mz, ppm):
  return(mz*(1-ppm/1000000), mz*(1+ppm/1000000))

conn = duckdb.connect("demo_data/msdata.duckdb")
pre_statement = "SELECT * FROM MS2 WHERE fragmz BETWEEN ? AND ?"
pre_data = conn.execute(pre_statement, pmppm(59.07336, 10)).fetchdf()
fig, axes = plt.subplots(3, 1, sharex=True)
for ax, filename in zip(axes, pre_data["filename"].unique()):
    subset = pre_data[pre_data["filename"] == filename]
    ax.vlines(subset["premz"], 0, subset["int"], color='k', zorder=1, linewidths=0.5)
    ax.scatter(subset["premz"], subset["int"], color='k', zorder=2, s=10)
    ax.set_yscale("log")
    ax.set_title(filename)
conn.close()
plt.tight_layout()
plt.show()
```

</details>

<details>
<summary>Julia</summary>

Make sure the SQLite/DuckDB, DataFrames, DBInterface, and StatsPlots packages are installed.

```{julia jfrag}
using DuckDB, DataFrames, StatsPlots;
pmppm(mass, ppm) = [mass * (1 - ppm / 1e6), mass * (1 + ppm / 1e6)];

conn = DBInterface.connect(DuckDB.DB, "demo_data/msdata.duckdb");
pre_statement = "SELECT * FROM MS2 WHERE fragmz BETWEEN ? AND ?";
pre_data = DataFrame(DBInterface.execute(conn, pre_statement, pmppm(59.07336, 10)));
DBInterface.close!(conn)

@df pre_data scatter(:premz, :int, markersize=3, group=:filename, yscale=:log10)
# @df pre_data plot!([:premz :premz]', [zeros(size(pre_data, 1)) pre_data.int]', lw=1, group=:filename)
```

</details>

## Other README things